Streaming CityJSON datasets

### 2024-04-19 paper/main.tex

Note that a \texttt{CityJSONFeature} object does not contain all the information that is required for parsing the feature. 
Most commonly, the \texttt{"transform"} property, the CRS, and the \texttt{"geometry-templates"} must be known by the client in order to correctly reconstruct and georeference the city objects. 
The rule \#4 ensures that those are available. 
The CityJSON object must contain \texttt{"transform"} and eventually the other properties if needed; \texttt{"CityObjects"} and \texttt{"vertices"} must be present but be empty (to ensure that the JSON object is valid).

@@

First observe that---contrary to intuition---the size of a dataset, serialised to a JSON file, stored as a CityJSONSeq is around 12\% compacter than as a CityJSON, and in the case \textbf{3DBAG} it is 36\% compacter.
The main reason is that the indices of the vertices are low integers for each feature (because the lowest index in each feature is always ``0''), and no not increase to several millions.
For instance, the dataset \textbf{Helsinki} contains more than 3 millions vertices, but its largest feature contains only but 2202.
As a result, the CityJSONSeq stored in a file is only but 72\% of the original CityJSON file.
Only one dataset sees its size slightly increase when serialised to a CityJSONSeq file (\textbf{Rotterdam}).
This is caused, we believe, by the fact that the dataset has textures that are shared by most of the buildings, and the \texttt{"appearance"} property must be repeated for each of the feature (in the CityJSON file it is only stored once and referenced by all buildings).

@@

With a CityJSONSeq file, we can read the file line by line, processing and discarding the city objects one by one.
As shown in the experiments below, this allows a very efficient operation in terms of both CPU and memory usage.
Operations that would highly benefit from this are subsetting, merging a citymodel.
Things like EPSG reassignment or metadata updates wouldn't even require to loop through the features, just to alter the first object in the file.

We processed all the datasets from Table~\ref{tab:datasets} with 2 simple Python scripts that loops through each city objects and their geometries and increments a global counter for the geometry type (\texttt{Solid} or \texttt{MultiSurface}).
The results for both the maximum memory footprint\footnote{The resident set size (RSS) was used, which is the portion of main memory occupied by the Python script.} and the time used is shown in Table~\ref{tab:ramtime}.
\begin{table*}
  \centering
  \caption{}
  \small
  \begin{tabular}
    {@{}lcrrcrrr@{}}\toprule
    &&  \multicolumn{2}{c}{\textbf{RAM used (MB)}} && \multicolumn{3}{c}{\textbf{time (s)}} \\ 
    \cmidrule{3-4} \cmidrule{6-8} 
     && CityJSON & CityJSONSeq && CityJSON & CityJSONSeq & diff \\
    \midrule
     \textbf{3DBAG}         &&   76.9 & 16.1  &&   0.10 & 0.07 & 1.4 \\
     \textbf{Helsinki}      && 3743.1 & 15.0  &&  13.39 & 2.74 & 4.9 \\
     \textbf{Helsinki\_tex} && 5004.8 & 19.1  &&  29.60 & 4.72 & 6.3 \\
     \textbf{Ingolstadt}    &&   65.5 & 21.3  &&   0.08 & 0.06 & 1.3 \\
     \textbf{Montréal}      &&   79.3 & 20.8  &&   0.11 & 0.07 & 1.6 \\
     \textbf{NYC}           &&  949.5 & 16.0  &&   1.78 & 0.70 & 2.5 \\
     \textbf{Railway}       &&   69.6 & 29.6  &&   0.09 & 0.07 & 1.3 \\
     \textbf{Rotterdam}     &&   42.4 & 14.6  &&   0.04 & 0.04 & 1.0 \\
     \textbf{Vienna}        &&   60.1 & 15.7  &&   0.06 & 0.05 & 1.2 \\
     \textbf{Zurich}        && 2793.1 & 16.3  &&   6.05 & 2.00 & 3.0 \\
    \bottomrule
  \end{tabular}%
  \label{tab:ramtime}
\end{table*}

The results indicate that there is a very significant benefit to using CityJSONSeq, compared to regular CityJSON files.
At least for operations that do not require analysing or calculating city objects that are close to each other.
Operations like calculating volumes, merging and subsetting files, finding city objects having specific attributes, etc.\ will all use significantly less memory and be significantly faster.

>>>>

Note that a \texttt{CityJSONFeature} object does not contain all the information that is required for parsing the feature. 
Most commonly, the \texttt{"transform"} property, the CRS, and the \texttt{"geometry-templates"} must be known by the client in order to correctly reconstruct and georeference the city objects. 
The rule \#4 ensures that those are available. 
The CityJSON object must contain \texttt{"transform"} and eventually the other properties if needed; \texttt{"CityObjects"} and \texttt{"vertices"} must be present but they must be empty (to ensure that the JSON object is valid).

@@

First observe that---contrary to intuition---the size of a dataset, serialised to a JSON file, stored as a CityJSONSeq is around 12\% more compact than as a CityJSON, and in the case of \textbf{3DBAG} it is 36\% compacter.
The main reason for this is that the indices of the vertices are low integers for each feature (because the lowest index in each feature is always ``0''), and no not increase to several millions.
%@Hugo: Sentence above not super clear. 
For instance, the dataset \textbf{Helsinki} contains more than 3 millions vertices, but its largest feature contains only but 2202.
As a result, the CityJSONSeq stored in a file is only but 72\% of the original CityJSON file.
Only one dataset sees its size slightly increased when serialised to a CityJSONSeq file (\textbf{Rotterdam}).
This is caused, we believe, by the fact that the dataset has textures that are shared by most of the buildings, and the \texttt{"appearance"} property must be repeated for each of the feature (in the CityJSON file it is only stored once and referenced by all buildings).

@@

With a CityJSONSeq file, we can read the file line by line, processing and discarding the city objects one by one.
As shown in the experiments below, this allows for a very efficient operation in terms of both CPU and memory usage.
Operations that would highly benefit from this are subsetting, merging a citymodel.
Things like EPSG reassignment or metadata updates wouldn't even require to loop through the features, just to alter the first object in the file.

We processed all the datasets from Table~\ref{tab:datasets} with 2 simple Python scripts that loop through the city objects and their geometries and increment a global counter for the geometry type (\texttt{Solid} or \texttt{MultiSurface}).
The results for both the maximum memory footprint\footnote{The resident set size (RSS) was used, which is the portion of main memory occupied by the Python script.} and the time used are shown in Table~\ref{tab:ramtime}.
\begin{table*}
  \centering
  \caption{}
  \small
  \begin{tabular}
    {@{}lcrrcrrr@{}}\toprule
    &&  \multicolumn{2}{c}{\textbf{RAM used (MB)}} && \multicolumn{3}{c}{\textbf{time (s)}} \\ 
    \cmidrule{3-4} \cmidrule{6-8} 
     && CityJSON & CityJSONSeq && CityJSON & CityJSONSeq & diff \\
    \midrule
     \textbf{3DBAG}         &&   76.9 & 16.1  &&   0.10 & 0.07 & 1.4 \\
     \textbf{Helsinki}      && 3743.1 & 15.0  &&  13.39 & 2.74 & 4.9 \\
     \textbf{Helsinki\_tex} && 5004.8 & 19.1  &&  29.60 & 4.72 & 6.3 \\
     \textbf{Ingolstadt}    &&   65.5 & 21.3  &&   0.08 & 0.06 & 1.3 \\
     \textbf{Montréal}      &&   79.3 & 20.8  &&   0.11 & 0.07 & 1.6 \\
     \textbf{NYC}           &&  949.5 & 16.0  &&   1.78 & 0.70 & 2.5 \\
     \textbf{Railway}       &&   69.6 & 29.6  &&   0.09 & 0.07 & 1.3 \\
     \textbf{Rotterdam}     &&   42.4 & 14.6  &&   0.04 & 0.04 & 1.0 \\
     \textbf{Vienna}        &&   60.1 & 15.7  &&   0.06 & 0.05 & 1.2 \\
     \textbf{Zurich}        && 2793.1 & 16.3  &&   6.05 & 2.00 & 3.0 \\
    \bottomrule
  \end{tabular}%
  \label{tab:ramtime}
\end{table*}

The results indicate that there is a very significant benefit to using CityJSONSeq, compared to regular CityJSON files, at least for operations that do not require analysing or calculating city objects that are close to each other.
Operations like calculating volumes, merging and subsetting files, finding city objects with specific attributes, etc.\ will all use significantly less memory and will be significantly faster.
